---

title: "Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)"
author: "Alyssa Pradhan"
output: github_document

---

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data", "burden-of-disease-cmnn.csv"), show_col_types = FALSE)

```

## Table of Estimates for CMNN Burden Over Time

DALY by year for each of Australasia, Cameroon and Malawi

```{r}
# Here render a table for the DALY burden over time for the three countries 

DALY3 <- as_tibble(data_cmnn)
DALY3 <- DALY3 %>% filter(Entity == "Australasia" | Entity == "Malawi" | Entity == "Cameroon")

DALY3 <- DALY3 %>% rename(
  Country = Entity,
  DALY = `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`
)

DALY3 <- DALY3 %>% select(-Code)

# You should pivot the data to show each country in a separate column.

DALY_wide <- DALY3 %>% pivot_wider(id_cols = Year,
                                   names_from = Country,
                                   values_from = DALY)

# Use kable() from the knitr package to render the table.

knitr::kable(DALY_wide)

```

## Plot Showing Trends in CMNN Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

ggplot(DALY3, aes(x = Year, y = DALY, color = Country)) +
  geom_line() +
  ggtitle("Daily adjusted life years from communicable diseases over time") +
  xlab("Time") +
  ylab("Daily adjusted life years")

```

## Summary of CMNN Burden Findings

This data demonstrates that in Malawi and Cameroon there was an initial increase in DALY in the late 1990s and early 2000s, possibly due to the HIV/AIDs epidemic. This has reduced since the 2010s. The DALYs lost in Australia from communicable disease has generally been low.

