---
title: "Non-Communicable Disease Burden (NCD)"
subtitle: "Trends: Austrat,Malawi & Camerron"
author: "Clement Trapence"
output: github_document
---


```{r echo=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)
```


```{r echo=FALSE ,message=F,warning=F}

data_ncd <- read_csv(here("data", "burden-of-disease-ncd.csv"))
```

## Table of Estimates for NCD Burden Over Time

```{r echo=FALSE}

data_ncd_clean <- data_ncd %>% 
  filter(Entity %in% c("Cameroon","Australia","Malawi")) %>% 
  rename(country=Entity,
         DALYs=`DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`,
         country_code=Code) 

data_ncd_clean %>% 
  kable()
```

## Plot Showing Trends in NCD Burden Over Time 

```{r}

data_ncd_clean %>% 
  ggplot(aes(x=Year,y=DALYs,group=country,
             color=country))+
  geom_line()+
  labs(title="DALYs Overtime- by country") +
  theme_classic()
```

## Summary of NCD Burden Findings

Australia (red line): There is a consistent downward trend in DALYs over the period. 
This suggests a decrease in the total disease burden and an improvement in population health over time.

Cameroon (green line): The DALYs for Cameroon appear to be relatively stable with a slight increase in the middle of the time period, followed by a modest decrease. 
This indicates that the disease burden has not changed dramatically over the years, with some fluctuation that could be due to specific health crises or changes in health policy or healthcare delivery.

Malawi (blue line): There's an increase in DALYs from around 1990 until around the year 2000, after which there is a steady decline. The initial increase could have been due to a worsening of health conditions or an outbreak of disease. The subsequent decline suggests an improvement in health conditions or effectiveness of health interventions.

```{r}

```



